Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving

### 2022-03-04 paper/main.tex
min tweak
experiments: progress

We start with several simple observations. Acceptance of a run $\pi$
is fully determined by the set $\minf(\pi)$.
% of acceptance marks that
% appear infinitely often on $\pi$, which is in turn fully determined by
% the set $\minf(\pi)$ of transitions appearing infinitely often in
% $\pi$.
Further, each set $\minf(\pi)$ for a run $\pi$ is a cycle and vice
versa. Hence, our reduction methods is looking for a new acceptance
formula $\psi$ and function $\rem$ such that for each cycle
$T\subseteq\delta$, it holds that $T$ satisfies $\varphi$ if and only
if $\rem(T)$ satisifes $\psi$. This can be roughly denoted by the
formula
\[
  \forall T\subseteq\delta~.~\mcycle(T)\implies\big(\msat_\varphi(T) ~\iff~ \msat_{\psi,\rem}(T)\big).
\]
In fact, this corresponds to the shape of the QBF formulas we will
construct. As we are looking for $\psi$ and $\rem$ such that the
formula holds, the subformula $\msat_{\psi,\rem}(T)$ contains many
free variables representing possible instances of $\psi$ and
$\rem$. If the formula is satisfiable, then its models encode desired
instances of $\psi$ and $\rem$.  In the following, we assume that we
are looking for a new acceptance formula $\psi$ in DNF. Note that our
reduction method can be easily adopted to look for $\psi$ of a
different shape.

@@

%\begin{table}[t]
%\centering
%\begin{tabular}{c|ccccc|}
%\multicolumn{1}{c}{} & 3 & 4 & 5 & 6 & \multicolumn{1}{c}{7} \\
%%\specialrule{2pt}{0pt}{0pt}
%\cline{2-6}
%3 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \cellcolor{blue!25}\makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
%4 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
%5 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
%6 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
%7 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
%\cline{2-6}
%\end{tabular}
%\end{table}

@@

>>>>

We start with several simple observations. Acceptance of a run $\pi$
is fully determined by the set $\minf(\pi)$.
% of acceptance marks that
% appear infinitely often on $\pi$, which is in turn fully determined by
% the set $\minf(\pi)$ of transitions appearing infinitely often in
% $\pi$.
Further, each set $\minf(\pi)$ for a run $\pi$ is a cycle and vice
versa. Hence, our reduction methods is looking for a new acceptance
formula $\psi$ and function $\rem$ such that for each cycle
$T\subseteq\delta$, it holds that $T$ satisfies $\varphi$ if and only
if $\rem(T)$ satisifes $\psi$. This can be roughly denoted by the
formula
\[
  \forall T\subseteq\delta~.~\mcycle(T)\implies\big(\msat_\varphi(T) ~\iff~ \msat_{\psi,\rem}(T)\big).
\]
In fact, this corresponds to the shape of the QBF formulas we will
construct. As we are looking for $\psi$ and $\rem$ such that the
formula holds, the subformula $\msat_{\psi,\rem}(T)$ contains many
free variables representing possible instances of $\psi$ and
$\rem$. If the formula is satisfiable, then its models encode desired
instances of $\psi$ and $\rem$.  In the following, we assume that we
are looking for a new acceptance formula $\psi$ in DNF. Note that our
reduction method can be easily adapted to look for $\psi$ of a
different shape.

@@

@@

Table~\ref{tab:heatmap} shows the number of automata coming from all four sets
for every input and output number of acceptance marks. The results of the
multi-level reduction technique are presented. The table indicates that in most
cases only 1 or 2 marks can be saved. However, some of the automata with
more than 7 acceptance marks could be, in theory, probably reduced more as
the query to the QBF solver mostly remains unanswered in the time limit.
Moreover, some of the results are promising, e.g. in 17 cases, we have
successfully reduced 8 or more acceptance marks to only 4.

\begin{table}[t]
\caption{The effect of reduction.
A cell on coordinates $(x,y)$ contains the number of automata that have been
reduced from $x$ to $y$ acceptance marks. If the cell contains a sum of two
numbers, the latter represents number of automata where the last query to QBF
solver has not been answered in the time limit.}
\label{tab:heatmap}
\setlength{\tabcolsep}{4pt}
\centering
\input{table-heatmap.tex}
\end{table}
